---
title: "Outlier Demonstration"
author: "Peiyuan Zhou"
date: "01/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5,
                      fig.path='Figs/', echo = TRUE)
```

This section is for the basic set up.
It will clear all the plots, the console and the workspace.
It also sets the overall format for numbers.

```{r}
if(!is.null(dev.list())) dev.off()
cat("\014") 
rm(list=ls())
options(scipen=9)
```
This section loads and attaches all the necessary packages.

```{r}
#Load packages

if(!require(readxl)){install.packages("readxl")}
library("readxl")

if(!require(pastecs)){install.packages("pastecs")}
library("pastecs")

if(!require(ggplot2)){install.packages("ggplot2")}
library("ggplot2")

if(!require(lattice)){install.packages("lattice")}
library("lattice")

if(!require(tinytex)){install.packages("tinytex")}
library("tinytex")
```

### Read in Data

```{r}
df <- read_excel("Outlier_Example.xlsx")
head(df)
df
```

### Box Plots for 1.5 IQR
```{r}
#As for all numerical data, check Box Plots
boxplot( mpg ~ cyl, data = mtcars,
         xlab = "No of Gallons",
         ylab = "Miles per Gallon",
         main = "Milege Data",
         notch = TRUE,
         varwidth = TRUE,
         col = c("green","yellow","purple"),
         names = c("High","Medium","Low"))

```
 
### Box Plots for 1.5 IQR
```{r}
#As for all numerical data, check Box Plots
boxplot(df$x1, horizontal=TRUE, pch=20)
#quantile(df$x1)
#a <- -0.62999-1.5*IQR(df$x1)
#b <- 0.68780+1.5*IQR(df$x1)

boxplot(df$x2, horizontal=TRUE, pch=20)
boxplot(df$x3, horizontal=TRUE, pch=20)
boxplot(df$x4, horizontal=TRUE, pch=20)
boxplot(df$x5, horizontal=TRUE, pch=20)
boxplot(df$x6, horizontal=TRUE, pch=20)
boxplot(df$x8, horizontal=TRUE, pch=20)
boxplot(df$x9, horizontal=TRUE, pch=20)

#Adjusting the IQR (for fun)
boxplot(df$x1, horizontal=TRUE, pch=20, range=2)
boxplot(df$x2, horizontal=TRUE, pch=20, range=2)
```
 
### Density Plots for further details
```{r}
densityplot( ~ df$x6, pch=6)
densityplot( ~ df$x2, pch=6)
densityplot( ~ df$x3, pch=6)
```
 Clearly 
 x2 has an anomolous value at ~ 10
 x3 has an anomolous value at ~ -10
 x6 has an odd value at ~ 0
 
 
### Scatter-Plots can help spot "hidden" outliers
```{r}
densityplot( ~ df$x8, pch=6, main='x8 looks good')
densityplot( ~ df$x9, pch=6, main='x9 looks good')
plot(df$x8,df$x9, main='Hunting Hiding outliers',pch=20)
```

### Don't forget Factor/categorical Variables
```{r}
df$x7 <- as.factor(df$x7)
barplot(table(df$x7), cex.names=.75,col='blue')
#talbe(df$x7): counts the occurrences of each unique value in x7
#cex.names: adjust the size of the axis labels (names) on the x-axis

#install.packages("DescTools")
#library(DescTools)
#Freq(df$x7)
```
Clearly Shanghai is a contextual outlier

## Code to Deal with outliers
### delete the row
```{r}
densityplot( ~ df$x6, pch=6)
nr <- which(df$x6 == 0)  #Find row number with x6 = 0
df <- df[-c(nr),]
densityplot( ~ df$x6, pch=6) # show the plot after removing the outlier.

densityplot( ~ df$x2, pch=6)
head(df[rev(order(df$x2)),],3)   #See the top three
nr <- which(df$x2 == max(df$x2))  #Row number for max value 
df <- df[-c(nr),]
densityplot( ~ df$x2, pch=6)
```
### adjust the value
```{r}
head(df)
densityplot( ~ df$x3, pch=6)
df$x3 <- ifelse(df$x3 == min(df$x3),mean(df$x3),df$x3)  #Replaces min value with the mean
densityplot( ~ df$x3, pch=6)
```


##### Correlation

```{r}
df <- dataset

df[,-c(dataset$)]
corr_mat <- round(cor(data),2) 
head(corr_mat)
```

